\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-13 Documentation/TearingModeTheory.tex
Finished error field penetration section

\begin{document}
\title{\bf Tearing Mode Dynamics in Tokamak Plasmas}
\author{R.~Fitzpatrick\,\footnote{rfitzp@utexas.edu}\\
Institute for Fusion Studies,\\  Department of Physics,\\  University of Texas at Austin,\\  Austin TX 78712, USA}
\maketitle

\section{Introduction}

@@

A tokamak plasma is {\em highly magnetized}. In other words, 
\begin{equation}
{\mit\Omega}_i\,\tau_i, \,\,\,{\mit\Omega}_e\, \tau_e \gg 1,
\end{equation}
 which implies that the electron and ion gyroradii are much  smaller than the
corresponding mean-free-paths between $90^\circ$ collisional scattering events. In this limit, a standard two-Laguerre-polynomial
Chapman-Enskog closure scheme (Chapman \& Cowling 1953) yields
\begin{align}\label{e3.91a}
{\bf F}&= n\,e\left(\frac{{\bf j}_\parallel}{\sigma_\parallel}
+\frac{{\bf j}_\perp}{\sigma_\perp}\right) -0.71\,n\,\nabla_\parallel T_e
-\frac{3\,n}{2\,{\mit\Omega}_e\,\tau_e}\,{\bf b}\times\nabla_\perp T_e,
\\[0.5ex]
w_i &= \frac{3\,m_e}{m_i} \frac{n\,(T_e-T_i)}{\tau_e},\\[0.5ex]
w_e &= -w_i + \frac{ {\bf j}\cdot {\bf F} }{n \,e}.
\end{align}
Here, ${\bf b}={\bf B}/B$ is a unit vector parallel to the magnetic field, and ${\bf j}= n\,e\,({\bf V}_i-{\bf V}_e)$ is the plasma current density. 
Moreover, the {\em parallel electrical conductivity}\/ is given by
\begin{equation}\label{e3.83}
\sigma_\parallel = 1.96\,\frac{n \,e^2\,\tau_e}{m_e}.
\end{equation} 
whereas the {\em perpendicular electrical conductivity}\/ takes the form
\begin{equation}\label{e3.92}
\sigma_\perp = 0.51\,\sigma_\parallel = \frac{n\,e^2\,\tau_e}{m_e}.
\end{equation}
Note that $\nabla_\parallel(\cdots) \equiv [{\bf b}\cdot\nabla
(\cdots)]\,{\bf b}$ denotes a
gradient parallel to the magnetic field, whereas $\nabla_\perp \equiv
 \nabla-\nabla_\parallel$ denotes a gradient perpendicular to the magnetic
field. Likewise, ${\bf j}_\parallel \equiv ({\bf b}\cdot{\bf j})\,{\bf b}$
represents the component of the plasma current density  flowing parallel to the
magnetic field, whereas ${\bf j}_\perp \equiv {\bf j} - {\bf j}_\parallel$
represents the perpendicular component of the plasma current density.

@@

In a highly magnetized plasma, the viscosity tensor is best described as the
sum of  five component tensors,
\begin{equation}
\bpi  = \sum_{n=0,4} \bpi_n,
\end{equation}
where 
\begin{equation}
\bpi_0 = - 3\,\eta_0\,\left({\bf b}{\bf b} - \frac{1}{3}\,{\bf I}\right)
\left({\bf b}{\bf b} - \frac{1}{3}\,{\bf I}\right): \nabla {\bf V},
\end{equation}
with
\begin{equation}
\bpi_1 =- \eta_1\left[{\bf I}_\perp \cdot{\bf W}\cdot{\bf I}_\perp
+ \frac{1}{2}\,{\bf I}_\perp\,({\bf b}\cdot{\bf W}\cdot{\bf b})\right],
\end{equation}
and
\begin{equation}
\bpi_2 = -4\,\eta_1\,\left({\bf I}_\perp\cdot
{\bf W}\cdot{\bf b}{\bf b}
+ {\bf b}{\bf b}\cdot{\bf W} \cdot{\bf I}_\perp\right).
\end{equation}
plus
\begin{equation}
\bpi_3 = \frac{\eta_3}{2}\,\left( {\bf b}\times 
{\bf W}\cdot{\bf I}_\perp - {\bf I}_\perp\cdot{\bf W}\times{\bf b}
\right),
\end{equation}
and
\begin{equation}
\bpi_4 = 2\,\eta_3\,\left({\bf b} \times{\bf W} \cdot {\bf b}{\bf b}
- {\bf b}{\bf b} \cdot{\bf W} \times{\bf b}\right).
\end{equation}
Here, ${\bf I}$ is the identity tensor, and 
${\bf I}_\perp = {\bf I} - {\bf b}{\bf b}$. The previous 
expressions are valid for both electrons and ions.

@@

Let us, first of all, consider the electron fluid equations, which can
be written:
\begin{align}
\frac{d_e n}{dt} + n\,\nabla\cdot{\bf V}_e &= 0,\\[0.5ex]
m_e\, n\,\frac{d_e {\bf V}_e}{dt} + \nabla p_e+ \nabla\cdot \bpi_e + e\, n\,
({\bf E} + {\bf V}_e\times {\bf B})& = {\bf F}_U +{\bf F}_T,\\[0.5ex]
\frac{3}{2}\frac{d_e p_e}{dt} + \frac{5}{2}\,p_e\,\nabla\cdot{\bf V}_e
+ \bpi_e:\nabla{\bf V}_e+ \nabla\cdot{\bf q}_{Te} 
+ \nabla\cdot{\bf q}_{Ue} &= -w_i+w_U + w_T.
\end{align}
Let $\bar{n}$, $\bar{v}_e$, $\bar{\tau}_e$, $\bar{l}_e=\bar{v}_e\,\bar{\tau}_e$, $\bar{B}$,
and $\bar{\rho}_e =\bar{ v}_e/(e\bar{B}/m_e)$,  be typical values
of the particle density, the electron thermal velocity, the electron collision time, the electron
mean-free-path between collisions,  the magnetic field-strength, and the
electron gyroradius,  respectively. 
Suppose that the typical spatial variation lengthscale of fluid variables is $L$. Let
 \begin{align}
\delta_e &=\frac{\bar{\rho}_e}{L},\\[0.5ex]
\zeta_e &= \frac{\bar{\rho}_e}{\bar{l}_e},\\[0.5ex]
\mu &= \sqrt{\frac{m_e}{m_i}}.
\end{align}
All three of these parameters are assumed to be  small
compared to unity. Finally, the typical electron flow velocity is assumed to be of order $\delta_e\,\bar{v}_e$. This corresponds
to the so-called {\em drift ordering}\/ in which the flow velocity is comparable to the curvature and grad-B particle
drift velocities (Fitzpatrick 2015). The drift ordering is appropriate to tearing modes in tokamak plasmas, which are comparatively
slowly growing instabilities.

@@

Let us now consider the ion fluid equations, which can be written:
\begin{align}
\frac{d_i n}{dt} + n\,\nabla\cdot{\bf V}_i &= 0,\\[0.5ex]
m_i\, n\,\frac{d_i {\bf V}_i}{dt} + \nabla p_i + \nabla\cdot \bpi_i - e\, n\,
({\bf E} + {\bf V}_i\times {\bf B})& =- {\bf F}_U -{\bf F}_T,\\[0.5ex]
\frac{3}{2}\frac{d_i p_i}{dt} + \frac{5}{2}\,p_i\,\nabla \cdot {\bf V}_i
+ \bpi_i:\nabla{\bf V}_i+ \nabla\cdot{\bf q}_i & =w_i.
\end{align}
It is convenient to adopt a normalization scheme for the ion equations
which is similar to, but independent of, that employed to normalize the
electron equations. Let  $\bar{n}$, $\bar{v}_i$, $\bar{l}_i$, $\bar{B}$,
and $\bar{\rho}_i =\bar{ v}_i/(e\bar{B}/m_i)$,  be typical values
of the particle density, the ion  thermal velocity, the ion
mean-free-path between collisions,  the magnetic field-strength, and the
ion gyroradius,  respectively. 
Suppose that 
the typical spatial variation lengthscale of fluid quantities is $L$. Let
 \begin{align}
\delta_i &= \frac{\bar{\rho}_i}{L},\\[0.5ex]
\zeta_i &= \frac{\bar{\rho}_i}{\bar{l}_i},\\[0.5ex]
\mu &= \sqrt{\frac{m_e}{m_i}}.
\end{align}
All three of these parameters are assumed to be  small 
compared to unity.  As before, we adopt the drift ordering in which the typical ion flow velocity is assumed to be of order $\delta_i\,\bar{v}_i$.

@@

Note that Equation~(\ref{e100}) is singular at the so-called {\em resonant}\/ magnetic flux-surface, radius $r=r_s$, at which 
\begin{equation}\label{e104}
q(r_s)= \frac{m}{n}.
\end{equation}
At the resonant surface, ${\bf k}\cdot{\bf B}=0$, where ${\bf k} = (0,\,m/r,\,-n/R_0)$ is the wavevector of the tearing perturbation. 
The value of $\psi(r,t)$ at the resonant surface---namely, ${\mit\Psi}_s(t)$---is known as the {\em reconnected magnetic flux}\/ (Fitzpatrick 1993). 
Note that ${\mit\Psi}_s$ is, in general, a complex quantity. 
Let $\rho=(r-r_s)/r_s$. 
The solution of Equation~(\ref{e100}) in the vicinity of the resonant surface is
\begin{equation}\label{e105}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s+}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho>0$, and
\begin{equation}\label{e106}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s-}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho<0$. Here, 
\begin{align}
\alpha_s& =- \left(\frac{r\,q\,J_z'}{s}\right)_{r=r_s},\\[0.5ex]
s(r) &= \frac{r\,q'}{q}.\label{shear}
\end{align}
Moreover,  the real parameters ${\mit\Delta}_{s+}$ and  ${\mit\Delta}_{s-}$ are fully determined by Equation~(\ref{e100})
and the boundary conditions (\ref{e101})--(\ref{e103}). Note that, in general, $\psi$ is continuous across the resonant surface (in accordance with Maxwell's equations),
whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial \psi/\partial r$ implies the presence of a radially-thin, helical,  current sheet  at the resonant surface. This current sheet can only be resolved by retaining more terms in the perturbed plasma equation of motion, (\ref{e96}). (See Section~\ref{linear}.)

The complex quantity
\begin{equation}\label{e109}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing at the resonant surface. Matching the solutions
in the so-called {\em inner region}\/ (i.e., the region of the plasma in the immediate vicinity of the resonant surface)
and the so-called {\em outer region}\/ (i.e., everywhere in the plasma other than the inner region) with the help
of Equations~(\ref{e99}), (\ref{e105}), and (\ref{e106}), we obtain
\begin{equation}\label{e110}
{\mit\Delta\Psi}_s = E_{ss}\,{\mit\Psi}_s,
\end{equation}
where $E_{ss}= {\mit\Delta}_{s+}-{\mit\Delta}_{s-}$ is a real quantity that is known as the {\em tearing stability index}\/ (Furth, et al.\ 1963).

\subsection{Solution in Presence of Resistive Wall}\label{resistive}
Suppose, now, that the wall at $r=r_w$ possesses a finite electrical resistivity, but is surrounded by a second perfectly-conducting wall
located at radius $r_c>r_w$.
The solution in the outer region can be written
\begin{equation}\label{e111}
\psi(r,t) = {\mit\Psi}_s(t)\,\hat{\psi}_s(r) + {\mit\Psi}_w(t)\,\hat{\psi}_w(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is specified in the previous section, and the real function  $\hat{\psi}_w(r)$ is a solution of 
\begin{equation}\label{e100a}
\frac{d^2\hat{\psi}_w}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_w}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_w - \frac{J_z'\,\hat{\psi}_w}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101a}
\hat{\psi}_w(r\leq r_s) &= 0,\\[0.5ex]
\hat{\psi}_w(r_w) &= 1,\label{e102a}\\[0.5ex]
\hat{\psi}_w(r\geq r_c) &= 0.\label{e103a}
\end{align}
It is easily seen that
\begin{equation}\label{e117}
\hat{\psi}_w(r_w< r < r_c) = \frac{(r/r_c)^{-m} - (r/r_c)^m}{(r_w/r_c)^{-m} - (r_w/r_c)^m}.
\end{equation}

@@

\subsection{Resistive Wall Physics}
It is clear from Equation~(\ref{e93}) that
\begin{equation}
\delta {\bf A} \simeq \psi\,{\bf e}_z,
\end{equation}
where $\delta {\bf A}$ is the perturbed magnetic vector potential. 
Hence, the perturbed electric field can be written
\begin{equation}
\delta {\bf E} \simeq -\nabla\delta{\mit\Phi} - \frac{\partial\psi}{\partial t}\,{\bf e}_z,
\end{equation}
where $\delta{\mit\Phi}$ is the perturbed electric scalar potential. 
Assuming that $|\nabla\delta{\mit\Phi}|\sim |\partial\psi/\partial t|$, the $z$-component of the previous
equation yields
\begin{equation}
\delta E_z\simeq -\frac{\partial\psi}{\partial t},
\end{equation}
where use has been made of the large aspect-ratio orderings (\ref{e92}). Within the wall, we
can write
\begin{equation}
\delta E_z \simeq -\frac{d{\mit\Psi}_w}{dt}.
\end{equation}
Here, we are making use of the so-called {\em thin wall approximation}, according to which $\psi$ is assumed to only vary weakly in $r$ across 
the wall. Ohm's law implies that
\begin{equation}
\delta j_z = \frac{\delta E_z}{\eta_w} = -\frac{1}{\eta_w}\,\frac{d{\mit\Psi}_w}{dt}
\end{equation}
within the wall, where $\eta_w$ is the wall resistivity. Equations~(\ref{e94}) and (\ref{e118}) yield
\begin{equation}\label{e132}
{\mit\Delta\Psi}_w = -\mu_0\int_{r_{w-}}^{r_{w+}}r\,\delta j_z\,dr.
\end{equation}
Suppose that the radial thickness of the wall is $\delta_w\ll r_w$. The previous two equations give
\begin{equation}\label{e133}
{\mit\Delta\Psi}_w = \tau_w\,\frac{d{\mit\Psi}_w}{dt},
\end{equation}
where
\begin{equation}
\tau_w = \frac{\mu_0\,r_w\,\delta_w}{\eta_w}
\end{equation}
is the so-called {\em time constant}\/ of the wall (Nave \& Wesson 1990; Fitzpatrick 1993). The thin wall approximation is
valid as long as
\begin{equation}
\frac{r_w}{\delta_w}\gg \tau_w\left|\frac{d\ln{\mit\Psi}_w}{dt}\right|.
\end{equation}

\subsection{Resonant Layer Physics}\label{sres}
By analogy with Equation~(\ref{e133}), we can write
\begin{equation}\label{e134}
{\mit\Delta\Psi}_s = \tau_s\left(\frac{d{\mit\Psi}_s}{dt} + {\rm i}\,\varpi_s\,{\mit\Psi}_s\right).
\end{equation}
Here, $\tau_s$ is the {\em reconnection time}\/ (i.e., the typical timescale on which magnetic reconnection takes place in the
resonant layer surrounding the resonant surface), whereas $\varpi_s$ is the so-called {\em natural frequency}. 
The natural frequency is the phase velocity of the tearing mode when the wall is perfectly conducting. The
natural frequency is non-zero because the reconnected flux in the resonant layer is convected by the local plasma
flow (Fitzpatrick 1993).

\subsection{Solution in Presence of External Magnetic Field-Coil}
Suppose that the perfectly-conducting wall at $r=r_c$ is replaced by a radially-thin, magnetic field-coil that carries a
helical current possessing $m$ periods in the poloidal direction, and $n$ periods in the toroidal direction. Let the current density in the field-coil
take the form
\begin{equation}\label{e135}
\delta j_z = \frac{I_c(t)}{r_c}\,\delta(r-r_c)\,{\rm e}^{\,{\rm i}\,(m\,\theta-n\,\varphi)}.
\end{equation}
Here, the complex quantity $I_c(t)$ parameterizes the amplitude and phase of the helical current flowing in the field-coil.

@@

Equations~(\ref{e220})--(\ref{e223}) are known collectively as the {\em drift-MHD equations}, and contain both shear-Alfv\'{e}n
and compressible-Alfv\'{e}n wave dynamics. The system of equations is completed by Maxwell's equations:
\begin{align}
\nabla\cdot{\bf B} &= 0,\\[0.5ex]
\nabla\times {\bf E} &= - \frac{\partial {\bf B}}{\partial t},\\[0.5ex]
\mu_0\,{\bf j} &= \nabla \times {\bf B}.\label{e226}
\end{align}

>>>>

\begin{document}
\title{\bf Tearing Mode Dynamics in Tokamak Plasmas}
\author{R.~Fitzpatrick\,\footnote{rfitzp@utexas.edu}\\
Institute for Fusion St udies,\\  Department of Physics,\\  University of Texas at Austin,\\  Austin TX 78712, USA}
\date{}
\maketitle

\tableofcontents

\newpage
\section{Introduction}

@@

A tokamak plasma is {\em highly magnetized}. In other words, 
\begin{equation}
{\mit\Omega}_i\,\tau_i, \,\,\,{\mit\Omega}_e\, \tau_e \gg 1,
\end{equation}
 which implies that the electron and ion gyroradii are much  smaller than the
corresponding mean-free-paths between $90^\circ$ collisional scattering events. In this limit, a standard two-Laguerre-polynomial
Chapman-Enskog closure scheme (Chapman \& Cowling 1953; Braginskii 1965) yields
\begin{align}\label{e3.91a}
{\bf F}&= n\,e\left(\frac{{\bf j}_\parallel}{\sigma_\parallel}
+\frac{{\bf j}_\perp}{\sigma_\perp}\right) -0.71\,n\,\nabla_\parallel T_e
-\frac{3\,n}{2\,{\mit\Omega}_e\,\tau_e}\,{\bf b}\times\nabla_\perp T_e,
\\[0.5ex]
w_i &= \frac{3\,m_e}{m_i} \frac{n\,(T_e-T_i)}{\tau_e},\\[0.5ex]
w_e &= -w_i + \frac{ {\bf j}\cdot {\bf F} }{n \,e}.
\end{align}
Here, ${\bf b}={\bf B}/B$ is a unit vector parallel to the magnetic field, and ${\bf j}= n\,e\,({\bf V}_i-{\bf V}_e)$ is the plasma current density. 
Moreover, the {\em parallel electrical conductivity}\/ is given by
\begin{equation}\label{e3.83}
\sigma_\parallel = 1.96\,\frac{n \,e^2\,\tau_e}{m_e}.
\end{equation} 
whereas the {\em perpendicular electrical conductivity}\/ takes the form
\begin{equation}\label{e3.92}
\sigma_\perp = 0.51\,\sigma_\parallel = \frac{n\,e^2\,\tau_e}{m_e}.
\end{equation}
Note that $\nabla_\parallel(\cdots) \equiv [{\bf b}\cdot\nabla
(\cdots)]\,{\bf b}$ denotes a
gradient parallel to the magnetic field, whereas $\nabla_\perp \equiv
 \nabla-\nabla_\parallel$ denotes a gradient perpendicular to the magnetic
field. Likewise, ${\bf j}_\parallel \equiv ({\bf b}\cdot{\bf j})\,{\bf b}$
represents the component of the plasma current density  flowing parallel to the
magnetic field, whereas ${\bf j}_\perp \equiv {\bf j} - {\bf j}_\parallel$
represents the perpendicular component of the plasma current density.

@@

In a highly magnetized plasma, the viscosity tensor is conveniently  described as the
sum of  five component tensors,
\begin{equation}
\bpi  = \sum_{n=0,4} \bpi_n,
\end{equation}
where 
\begin{equation}
\bpi_0 = - 3\,\eta_0\,\left({\bf b}{\bf b} - \frac{1}{3}\,{\bf I}\right)
\left({\bf b}{\bf b} - \frac{1}{3}\,{\bf I}\right): \nabla {\bf V},
\end{equation}
with
\begin{equation}
\bpi_1 =- \eta_1\left[{\bf I}_\perp \cdot{\bf W}\cdot{\bf I}_\perp
+ \frac{1}{2}\,{\bf I}_\perp\,({\bf b}\cdot{\bf W}\cdot{\bf b})\right],
\end{equation}
and
\begin{equation}
\bpi_2 = -4\,\eta_1\,\left({\bf I}_\perp\cdot
{\bf W}\cdot{\bf b}{\bf b}
+ {\bf b}{\bf b}\cdot{\bf W} \cdot{\bf I}_\perp\right).
\end{equation}
plus
\begin{equation}
\bpi_3 = \frac{\eta_3}{2}\,\left( {\bf b}\times 
{\bf W}\cdot{\bf I}_\perp - {\bf I}_\perp\cdot{\bf W}\times{\bf b}
\right),
\end{equation}
and
\begin{equation}
\bpi_4 = 2\,\eta_3\,\left({\bf b} \times{\bf W} \cdot {\bf b}{\bf b}
- {\bf b}{\bf b} \cdot{\bf W} \times{\bf b}\right).
\end{equation}
Here, ${\bf I}$ is the identity tensor, and 
${\bf I}_\perp = {\bf I} - {\bf b}{\bf b}$. The previous 
expressions are valid for both electrons and ions.

@@

Let us, first of all, consider the electron fluid equations, which can
be written:
\begin{align}
\frac{d_e n}{dt} + n\,\nabla\cdot{\bf V}_e &= 0,\\[0.5ex]
m_e\, n\,\frac{d_e {\bf V}_e}{dt} + \nabla p_e+ \nabla\cdot \bpi_e + e\, n\,
({\bf E} + {\bf V}_e\times {\bf B})& = {\bf F}_U +{\bf F}_T,\\[0.5ex]
\frac{3}{2}\frac{d_e p_e}{dt} + \frac{5}{2}\,p_e\,\nabla\cdot{\bf V}_e
+ \bpi_e:\nabla{\bf V}_e+ \nabla\cdot{\bf q}_{Te} 
+ \nabla\cdot{\bf q}_{Ue} &= -w_i+w_U + w_T.
\end{align}
Let $\bar{n}$, $\bar{v}_e$, $\bar{\tau}_e$, $\bar{l}_e=\bar{v}_e\,\bar{\tau}_e$, $\bar{B}$,
and $\bar{\rho}_e =\bar{ v}_e/(e\bar{B}/m_e)$  be typical values
of the particle density, the electron thermal velocity, the electron collision time, the electron
mean-free-path between collisions,  the magnetic field-strength, and the
electron gyroradius,  respectively. 
Suppose that the typical spatial variation lengthscale of fluid variables is $L$. Let
 \begin{align}
\delta_e &=\frac{\bar{\rho}_e}{L},\\[0.5ex]
\zeta_e &= \frac{\bar{\rho}_e}{\bar{l}_e},\\[0.5ex]
\mu &= \sqrt{\frac{m_e}{m_i}}.
\end{align}
All three of these parameters are assumed to be  small
compared to unity. Finally, the typical electron flow velocity is assumed to be of order $\delta_e\,\bar{v}_e$. This corresponds
to the so-called {\em drift ordering}\/ in which the flow velocity is comparable to the curvature and grad-B particle
drift velocities (Fitzpatrick 2015). The drift ordering is appropriate to tearing modes in tokamak plasmas, which are comparatively
slowly growing instabilities.

@@

Let us now consider the ion fluid equations, which can be written:
\begin{align}
\frac{d_i n}{dt} + n\,\nabla\cdot{\bf V}_i &= 0,\\[0.5ex]
m_i\, n\,\frac{d_i {\bf V}_i}{dt} + \nabla p_i + \nabla\cdot \bpi_i - e\, n\,
({\bf E} + {\bf V}_i\times {\bf B})& =- {\bf F}_U -{\bf F}_T,\\[0.5ex]
\frac{3}{2}\frac{d_i p_i}{dt} + \frac{5}{2}\,p_i\,\nabla \cdot {\bf V}_i
+ \bpi_i:\nabla{\bf V}_i+ \nabla\cdot{\bf q}_i & =w_i.
\end{align}
It is convenient to adopt a normalization scheme for the ion equations
which is similar to, but independent of, that employed to normalize the
electron equations. Let  $\bar{n}$, $\bar{v}_i$, $\bar{l}_i$, $\bar{B}$,
and $\bar{\rho}_i =\bar{ v}_i/(e\bar{B}/m_i)$  be typical values
of the particle density, the ion  thermal velocity, the ion
mean-free-path between collisions,  the magnetic field-strength, and the
ion gyroradius,  respectively. 
Suppose that 
the typical spatial variation lengthscale of fluid quantities is $L$. Let
 \begin{align}
\delta_i &= \frac{\bar{\rho}_i}{L},\\[0.5ex]
\zeta_i &= \frac{\bar{\rho}_i}{\bar{l}_i},\\[0.5ex]
\mu &= \sqrt{\frac{m_e}{m_i}}.
\end{align}
All three of these parameters are assumed to be  small 
compared to unity.  As before, we adopt the drift ordering in which the typical ion flow velocity is assumed to be of order $\delta_i\,\bar{v}_i$.

@@

Note that Equation~(\ref{e100}) is singular at the so-called {\em resonant}\/ magnetic flux-surface, radius $r=r_s$, at which 
\begin{equation}\label{e104}
q(r_s)= \frac{m}{n}.
\end{equation}
At the resonant surface, ${\bf k}\cdot{\bf B}=0$, where ${\bf k} = (0,\,m/r,\,-n/R_0)$ is the wavevector of the tearing perturbation. 
The value of $\psi(r,t)$ at the resonant surface---namely, ${\mit\Psi}_s(t)$---is known as the {\em reconnected magnetic flux}\/ (Fitzpatrick 1993). (See Section~\ref{dis1}.)
Note that ${\mit\Psi}_s$ is, in general, a complex quantity. 
Let $\rho=(r-r_s)/r_s$. 
The solution of Equation~(\ref{e100}) in the vicinity of the resonant surface is
\begin{equation}\label{e105}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s+}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho>0$, and
\begin{equation}\label{e106}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s-}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho<0$. Here, 
\begin{align}
\alpha_s& =- \left(\frac{r\,q\,J_z'}{s}\right)_{r=r_s},\\[0.5ex]
s(r) &= \frac{r\,q'}{q}.\label{shear}
\end{align}
Moreover,  the real parameters ${\mit\Delta}_{s+}$ and  ${\mit\Delta}_{s-}$ are fully determined by Equation~(\ref{e100})
and the boundary conditions (\ref{e101})--(\ref{e103}). Note that, in general, $\psi$ is continuous across the resonant surface (in accordance with Maxwell's equations),
whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial \psi/\partial r$ implies the presence of a radially-thin, helical,  current sheet  at the resonant surface. This current sheet can only be resolved by retaining more terms in the perturbed plasma equation of motion, (\ref{e96}). (See Section~\ref{linear}.)

The complex quantity
\begin{equation}\label{e109}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing at the resonant surface. Matching the solutions
in the so-called {\em inner region}\/ (i.e., the region of the plasma in the immediate vicinity of the resonant surface)
and the so-called {\em outer region}\/ (i.e., everywhere in the plasma other than the inner region) with the help
of Equations~(\ref{e99}), (\ref{e105}), and (\ref{e106}), we obtain
\begin{equation}\label{e110}
{\mit\Delta\Psi}_s = E_{ss}\,{\mit\Psi}_s,
\end{equation}
where $E_{ss}= {\mit\Delta}_{s+}-{\mit\Delta}_{s-}$ is a real dimensionless quantity that is known as the {\em tearing stability index}\/ (Furth, et al.\ 1963).

\subsection{Solution in Presence of Resistive Wall}\label{resistive}
Suppose, now, that the wall at $r=r_w$ possesses a finite electrical resistivity, but is surrounded by a second perfectly-conducting wall
located at radius $r_c>r_w$.
The solution in the outer region can be written
\begin{equation}\label{e111}
\psi(r,t) = {\mit\Psi}_s(t)\,\hat{\psi}_s(r) + {\mit\Psi}_w(t)\,\hat{\psi}_w(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is specified in Section~\ref{perfect}, and the real function  $\hat{\psi}_w(r)$ is a solution of 
\begin{equation}\label{e100a}
\frac{d^2\hat{\psi}_w}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_w}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_w - \frac{J_z'\,\hat{\psi}_w}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101a}
\hat{\psi}_w(r\leq r_s) &= 0,\\[0.5ex]
\hat{\psi}_w(r_w) &= 1,\label{e102a}\\[0.5ex]
\hat{\psi}_w(r\geq r_c) &= 0.\label{e103a}
\end{align}
It is easily seen that
\begin{equation}\label{e117}
\hat{\psi}_w(r_w< r < r_c) = \frac{(r/r_c)^{-m} - (r/r_c)^m}{(r_w/r_c)^{-m} - (r_w/r_c)^m}.
\end{equation}

@@

\subsection{Resistive Wall Physics}\label{tauw}
It is clear from Equation~(\ref{e93}) that
\begin{equation}
\delta {\bf A} \simeq \psi\,{\bf e}_z,
\end{equation}
where $\delta {\bf A}$ is the perturbed magnetic vector potential. 
Hence, the perturbed electric field can be written
\begin{equation}
\delta {\bf E} \simeq -\nabla\delta{\mit\Phi} - \frac{\partial\psi}{\partial t}\,{\bf e}_z,
\end{equation}
where $\delta{\mit\Phi}$ is the perturbed electric scalar potential. 
Assuming that $|\nabla\delta{\mit\Phi}|\sim |\partial\psi/\partial t|$, the $z$-component of the previous
equation yields
\begin{equation}
\delta E_z\simeq -\frac{\partial\psi}{\partial t},
\end{equation}
where use has been made of the large aspect-ratio orderings (\ref{e92}). Within the wall, we
can write
\begin{equation}
\delta E_z \simeq -\frac{d{\mit\Psi}_w}{dt}.
\end{equation}
Here, we are making use of the so-called {\em thin wall approximation}, according to which $\psi$ is assumed to only vary weakly in $r$ across 
the wall. Ohm's law implies that
\begin{equation}
\delta j_z = \frac{\delta E_z}{\eta_w} = -\frac{1}{\eta_w}\,\frac{d{\mit\Psi}_w}{dt}
\end{equation}
within the wall, where $\eta_w$ is the wall resistivity. Equations~(\ref{e94}) and (\ref{e118}) yield
\begin{equation}\label{e132}
{\mit\Delta\Psi}_w = -\mu_0\int_{r_{w-}}^{r_{w+}}r\,\delta j_z\,dr.
\end{equation}
Suppose that the radial thickness of the wall is $\delta_w\ll r_w$. The previous two equations give
\begin{equation}\label{e133}
{\mit\Delta\Psi}_w = \tau_w\,\frac{d{\mit\Psi}_w}{dt},
\end{equation}
where
\begin{equation}
\tau_w = \frac{\mu_0\,r_w\,\delta_w}{\eta_w}
\end{equation}
is the so-called {\em time constant}\/ of the wall (Nave \& Wesson 1990; Fitzpatrick 1993). The thin wall approximation is
valid as long as
\begin{equation}
\frac{r_w}{\delta_w}\gg \tau_w\left|\frac{d\ln{\mit\Psi}_w}{dt}\right|.
\end{equation}

\subsection{Resonant Layer Physics}\label{sres}
By analogy with Equation~(\ref{e133}), we can write
\begin{equation}\label{e134}
{\mit\Delta\Psi}_s = \tau_s\left(\frac{d{\mit\Psi}_s}{dt} + {\rm i}\,\omega_s\,{\mit\Psi}_s\right).
\end{equation}
Here, $\tau_s$ is the {\em reconnection time}\/ (i.e., the typical timescale on which magnetic reconnection takes place in the
resonant layer surrounding the resonant surface), whereas $\omega_s$ is the so-called {\em natural frequency}. 
The natural frequency is the phase velocity of the tearing mode when the wall is perfectly conducting. The
natural frequency is non-zero because the reconnected flux in the resonant layer is convected by the local plasma
flow (Fitzpatrick 1993). (See Section~\ref{s6.2}.)

\subsection{Solution in Presence of External Magnetic Field-Coil}\label{sfc}
Suppose that the perfectly-conducting wall at $r=r_c$ is replaced by a radially-thin, magnetic field-coil that carries a
helical current possessing $m$ periods in the poloidal direction, and $n$ periods in the toroidal direction. Let the current density in the field-coil
take the form
\begin{equation}\label{e135}
\delta j_z = \frac{I_c(t)}{r_c}\,\delta(r-r_c)\,{\rm e}^{\,{\rm i}\,(m\,\theta-n\,\varphi)}.
\end{equation}
Here, the complex quantity $I_c(t)$ parameterizes the amplitude and phase of the helical current flowing in the field-coil.

@@

Equations~(\ref{e220})--(\ref{e223}) are known collectively as the {\em drift-MHD equations},\footnote{The 'drift' appellation  merely indicates that we are taking diamagnetic flows into account.} and contain both shear-Alfv\'{e}n
and compressible-Alfv\'{e}n wave dynamics. The system of equations is completed by a subset of Maxwell's equations:
\begin{align}
\nabla\cdot{\bf B} &= 0,\\[0.5ex]
\nabla\times {\bf E} &= - \frac{\partial {\bf B}}{\partial t},\\[0.5ex]
\mu_0\,{\bf j} &= \nabla \times {\bf B}.\label{e226}
\end{align}

\subsection{Normalization Scheme}
Let $l$ be a typical variation lengthscale in the resonant layer. The {\em Alfv\'{e}n velocity}, which is the typical phase
velocity of compressible Alfv\'{e}n waves, is defined
\begin{equation}\label{e226a}
V_A = \frac{B_z}{\sqrt{\mu_0\,n_0\,m_i}}.
\end{equation}
It is helpful to define the  {\em collisionless ion skin-depth}: 
\begin{equation}
d_i = \left(\frac{m_i}{n_0\,e^{\,2}\,\mu_0}\right)^{1/2}.
\end{equation}
Note that $d_i=c/\omega_{p\,i}$, where $c$ is the velocity of light in vacuum, and $\omega_{p\,i}$ is the ion plasma
frequency (Fitzpatrick 2015).
